### 2019-10-30 SupplementaryMaterialECDModel.Rmd
start of summarizing analytical expressions for eco-evo dynamics (following Matt Osmond's suggestion).

---
title: "Supplementary Material"
author: "Matt Barbour"
date: "6/18/2019"
output:
  pdf_document:
    keep_tex: true
    toc: false
    number_sections: false
    df_print: paged
    fig_caption: true
    citation_package: natbib 
    includes:  
      in_header: Am-Nat-preamble-latex.tex
bibliography: references
biblio-style: amnat
fontsize: 11pt
fontfamily: mathpazo
documentclass: article
linkcolor: black
urlcolor: black
citecolor: black
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Resources occur in same habitat

## Assuming symmetry

In Appendix X, we show that in the absence of competition, consumer $i$ will reach an evolutionary stable strategy when:

$$a_{i,1}=a_{i,2}$$
This indicates that consumers will evolve to be complete generalists.

In the presence of a competitor though, the evolution of attack rates depends on the shape of the tradeoff. 
Specifically, when $n=1$, then the initial attack rates determine the ESS; therefore, the magnitude of character displacement depends on the initial attack rates.
When $n<1$, then consumers will evolve to be completely specialized on the resource that they are pre-adapted to. 
For example, if initial $$a_{i,1}>a_{i,2}$$, then consumer $i$ will evolve until $2a_{i,1}$ and $a_{i,2}=0$, and thus maximal character displacement.
In contrast, when $n>1$, then consumers will evolve to be complete generalists, $a_{i,1}=a_{i,2}$, and thus there will be no character displacement.

In the absence of competition, the ESS in attack rates results in equal resource abundances, which are determined by:

$$R=\frac{m}{2a_{i,1}e}$$
The effects of character displacement on resource abundances depends on the shape of the tradeoff. 
When $n\geq1$, then resource abundances are equivalent for both one and two consumer communities. 
However, when $n<1$, then

# Resources occur in different habitats

## Evolutionary Stable Strategy

In Appendix X, we show that a consumer (we arbitrarily chose $C_1$) will reach an evolutionary stable strategy when:

$$e_{11}R_1w_{11}\delta_{11} + e_{12}R_2w_{12}\delta_{12}=0$$

If there were no constraints, consumers would evolve higher attack rates on both resources. 
In consideration of this, we assume there is a trade-off in attack rates so that $\delta_{12}$ is a negative function of $delta_{11}$, $\delta_{12}=-f(\delta_{11})$.
Therefore, we can rewrite the above expression as

$$e_{11}R_1w_{11}\delta_{11} = e_{12}R_2w_{12}f(\delta_{11})$$

## Resource equilibrium

One consumer:

Two consumers:

## Local stability

One consumer:

Two consumers:

# Adding a more realistic functional response